---
title: "Using GeoDeepDive (xDD) to Track North American Bird Banding Laboratory Citations"
date: "Last updated: `r Sys.Date()`"
output: github_document
---
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-lightgrey.svg)](https://www.tidyverse.org/lifecycle/#experimental)

## About
This project seeks to identify publications within which the [USGS North American  Bird Banding Laboratory](https://www.usgs.gov/centers/pwrc/science/bird-banding-laboratory) has been referenced or mentioned. This project uses pre-extracted information available via API from the [GeoDeepDive](https://geodeepdive.org) infrastructure.

## Overview of BBL in records available in the GeoDeepDive database.

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
## if you open the .rproj you should be in the right directory, otherwise please set it here.
# setwd("path/to/directory/)

# source the R functions
source("R/get_xdd_df.R")


# load packages
  library(ggpubr)
  library(tidyverse)
  library(jsonlite)
  library(rjson)

# plotting themes and settings
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme_set(ggpubr::theme_classic2())
```

```{r bbl_xdd_import_and_munge, message=FALSE, warning=FALSE,echo=FALSE}
# Import the JSON files and munge into a single data frame ----------------
files <- list.files(here::here("xdd_json"), pattern='.txt', full.names=TRUE)

# import and munge the xdd json into an R df
bbl_xdd <- suppressWarnings(get_xdd_df(files)) # do not turn off suppresswarnings -- there an rmd bug its annoying..

# try to remove products related to oil and gas
bbl_xdd_no_oilgas <-  bbl_xdd %>% 
  filter(!str_detect(tolower(highlight), "gas|oil" ))

# remove the records with search term BBL:
bbl_xdd_reduced <- bbl_xdd %>% 
  filter(searchterm !="BBL")
```

We can use various search terms to identify potential uses of the BBL data or results in the literature. For example, here we found `r bbl_xdd %>% dplyr::select(-filename, -searchterm, -highlight) %>% n_distinct()` unique publications which used any of the following phrases exactly. However, "BBL" is a common acronym in oil and gas and other fields. We can coarsely try to control for this by removing references to "oil" or "gas" in the highlights--this results in `r bbl_xdd_no_oilgas %>% dplyr::select(-filename, -searchterm, -highlight) %>% n_distinct()` unqiue records -- this still leaves us with a lot of irrelevant publications. For our purposes, we will work only with the `r bbl_xdd_reduced %>% dplyr::select(-filename, -searchterm, -highlight) %>% n_distinct()` records associated with the following search terms:

Exact search phrase | Number of unique records in xDD/GDD 
--------------|----
"North American Bird Banding Laboratory" | `r bbl_xdd_reduced %>% filter(searchterm=="North American Bird Banding Laboratory") %>% dplyr::select(-filename, -highlight) %>% n_distinct()` 
<!-- "BBL" |  `r bbl_xdd_no_oilgas %>% filter(searchterm=="BBL") %>% dplyr::select(-filename, -highlight) %>% n_distinct()` -->
"Bird Banding Laboratory" | `r bbl_xdd_reduced %>% filter(searchterm=="Bird Banding Laboratory") %>% dplyr::select(-filename, -highlight) %>% n_distinct()` 


Check out the use of the terms over time: 
```{r pubsvtime, echo=FALSE}
tmp <- bbl_xdd_reduced %>% distinct(coveryear, pubname, publisher, title, authors, searchterm)   
ggplot(tmp %>% group_by(searchterm, coveryear) %>% 
         summarise(total_pubs = n()) %>% na.omit()) +
  geom_line(aes(x=coveryear, y=total_pubs, color=searchterm))+
  scale_color_manual(values=cbPalette)+
  labs(x="year", y="total records")+
  theme(legend.title=element_blank(), legend.position = "top")
``` 

We can visualize the distribution of  records among unique publishers. Few publications contain the bulk of the records. 
```{r topjrnls, echo=FALSE, message=FALSE, warning=FALSE}
tmp <-  bbl_xdd_reduced %>% dplyr::select(-filename, -searchterm) %>% distinct()
tmp <- bind_rows(tmp, bbl_xdd_reduced %>% dplyr::select(-filename, -searchterm)%>% filter(is.na(doi)) %>% distinct())

ggplot(tmp %>% group_by(pubname) %>% 
         summarise(n = n())) +
  geom_histogram(aes(n))+
  labs(x="number of publishers", y="total records")
```

Viewing the publications with at least 25 unique records:
```{r topjrnls2, echo=FALSE, message=FALSE, warning=FALSE}
tmp <- bbl_xdd_reduced %>% distinct(coveryear, pubname, publisher, title, authors, searchterm)%>% group_by(pubname) %>% 
         summarise(n = n())   

ggdotchart(
           tmp, x = "pubname", y = "n",
          rotate = TRUE,
          ylab="number of records", 
          xlab="",
            sorting = "desc",# title = "1980-present",
          ggtheme = theme_minimal()
          )
```


## This project relies on the UW Deepdive Infrastructure
[Original branch from which this repository was forked](UW-Deepdive-Infrastructure/app-template). Check out [the wiki](https://github.com/UW-Deepdive-Infrastructure/app-template/wiki) for more information on getting started.

## USGS Provisional Software
This software is preliminary or provisional and is subject to revision. It is being provided to meet the need for timely best science. The software has not received final approval by the U.S. Geological Survey (USGS). No warranty, expressed or implied, is made by the USGS or the U.S. Government as to the functionality of the software and related material nor shall the fact of release constitute any such warranty. The software is provided on the condition that neither the USGS nor the U.S. Government shall be held liable for any damages resulting from the authorized or unauthorized use of the software.
